### 2017-02-14 CO750-ApproximationAlgorithmsTopics/assignments/assignment_01.tex
conv opt a1 p3
fixed 4

\lhead{2017-02-09}
\rhead{William Justin Toth CO750-Approximation Algorithms Assignment 1} %CHANGE n to ASSIGNMENT NUMBER ijk TO COURSE CODE
\renewcommand{\headrulewidth}{1pt} %heading underlined
%\renewcommand{\baselinestretch}{1.2} % 1.2 line spacing for legibility (optional)

@@

It remains to verify the approximation factor of $O(1)$. We do so by induction on the number of iterations the algorithm needs to run to return a solution. In the base case, that is when $\delta = 0$, the algorithm returns an $O(1)$-approximate Steiner Tree connecting $R \cup \{r\}$ and installs capacity of $1$ on each edge. By observing that the optimal solution in this case is a Steiner Tree connecting $R \cup \{r\}$ with capacity $1$ on each edge we see that the approximation factor holds in this case.
Suppose that $\delta > 0$ and for any problem instance with maximum demand $2^{\alpha}$ with $0 \leq \alpha < \delta$ the algorithm returns an $O(1)$ approximate solution. We will describe solutions by an ordered pair $(T, x)$ where $T$ is a tree and $x : T \rightarrow \R_+$ maps edges of $T$ with their corresponding capacity. Then the objective function, which we denote by $f$, in our problem maps feasible solutions $(T,x)$ as follows:

@@

Let $(T,x)$ be the solution returned by running our algorithm, and let $(T^*, x^*)$ be the optimal solution. Let $T_{\delta}$ be the $O(1)$-approximate Steiner Tree connecting $R_\delta \cup \{r\}$ in the first iteration (iteration $\delta$) of our algorithm. Let $(\bar{T}, x)$ denote the solution returned by our algorithm running on the residual problem with input graph $G_{\delta-1}$ (the resulting graph after contracting $T_{\delta}$ into $r$). Then $T = T_\delta \cup \bar{T}$ and 
$$f(T,x) = f(T_\delta, x) + f(\bar{T},x).$$
If we let $(\bar{T}^*, \bar{x}^*)$ denote the optimal solution for the residual problem with graph $G_{\delta - 1}$ then by our inductive hypothesis
$$f(T,x) \leq f(T_\delta,x) + O(1)f(\bar{T}^*, \bar{x}^*).$$
Now we observe that $T^*$ contains a Steiner Tree $T'$ with terminals $R_\delta \cup \{r\}$. Since $T_\delta$ is an $O(1)$-approximate optimal Steiner tree with terminals $R_\delta \cup \{r\}$ we have $c(T_\delta) \leq O(1) c(T')$. Since $T'$ necessarily installs capacities $2^\delta$ on all of its edges, $x_e^* = 2^\delta$ for all $e \in T'$. To see that a capacity of $2^\delta$ is necessary, suppose some edge $e$ in $T'$ has $x^*_e < 2^\delta$. Since $T'$ is a Steiner Tree, some terminal $v \in R_\delta$ uses $e$ on its unique path to $r$. But then an $rv$-flow along this path has value at most $x^*_e < 2^\delta$ contradicting feasibility of $(T^*, x^*)$. Thus $x_e^* = 2^\delta$ for all $e \in T'$. Hence we have
$$f(T_\delta, x) = \sum_{e \in T_\delta} c_e x_e = 2^\delta c(T_\delta) \leq O(1)2^\delta c(T') = O(1)f(T', x^*) \leq O(1) f(T^*, x^*).$$
$$f(T,x) \leq O(1) f(T^*,x^*) + O(1) f(\bar{T}^*, \bar{x}^*).$$
Now consider the graph that results from contracting all edges of $T_\delta \cap T^*$ into $r$ in the tree $T^*$. If any cycles form then do the following: while there still exists a cycle $C$ take the edge $e = \text{argmin}_{e\in C} x^*_e$ and remove it. Call the resulting graph $S^*$. 

@@

We claim that $(S^*, x^*)$ is feasible for the residual problem on $G_{\delta-1}$. Indeed $S^*$ is a tree since we trimmed one edge from all cycles. Since trimming cycles preserves connectivity, and we contracted edges into $r$, all $rv$-paths in $T^*$ are preserved in $S^*$. Hence for all $v \in R\backslash R_\delta$ there exists an $rv$-path in $S^*$. Now we claim that the capacities on any $rv$-path still admit feasible flows after converting from $T^*$ to $S^*$. Observe that contraction will not decrease the minimum capacity. Further observe that deleted edges had minimum capacity on a cycle, so any flow that used a deleted edge $uv$ could send the same flow through the $uv$-path remaining from the cycle $uv$ was deleted from. Hence the flows on $S^*$ under capacities $x^*$ can still meet all demands in $R\backslash R_\delta$. Therefore $(S^*,x^*)$ (with $x^*$ restricted to $E(G_{\delta-1})$) is feasible for $G_{\delta -1}$. Hence, since $(\bar{T}^*, \bar{x}^*)$ is optimal for $G_{\delta -1}$:
$$f(\bar{T}^*, \bar{x}^*) \leq f(S^*, x^*) \leq f(T^*, x^*).$$
Therefore

@@

f(T,x) &\leq O(1)f(T^*,x^*) + O(1)f(\bar{T}^*,\bar{x}^*) \\
&\leq O(1)f(T^*,x^*) + O(1)f(T^*,x^*) \\
&= 2O(1)f(T^*,x^*)\\
&= O(1)f(T^*,x^*).

@@

So the approximation factor $O(1)$ holds in the inductive case. Thus the algorithm given is a $O(1)$-approximation algorithm.$\blacksquare$

>>>>

\lhead{2017-02-09}
\rhead{W. Justin Toth CO750-Approximation Algorithms Assignment 1} %CHANGE n to ASSIGNMENT NUMBER ijk TO COURSE CODE
\renewcommand{\headrulewidth}{1pt} %heading underlined
%\renewcommand{\baselinestretch}{1.2} % 1.2 line spacing for legibility (optional)

@@

It remains to verify the approximation factor of $O(1)$.We will describe solutions by an ordered pair $(T, x)$ where $T$ is a tree and $x : T \rightarrow \R_+$ maps edges of $T$ with their corresponding capacity. Then the objective function, which we denote by $f$, in our problem maps feasible solutions $(T,x)$ as follows:

@@

Let $(T,x)$ be the solution returned by running our algorithm, and let $(T^*, x^*)$ be the optimal solution. Let $T_{i}$ be the $O(1)$-approximate Steiner Tree connecting $R_i \cup \{r\}$ in iteration $i$ of our algorithm. Similarly let $T^*_i$ denote the sub-tree of $T^*$ which spans $\{r\} \cup R_i$ after contracting $T^*_{i+1} \dots T^*_\delta$. Observe that the capacities on edges, $e$ of $T_i$ or $T^*_i$ are necessarily $2^i$, as otherwise a flow of value $2^i$ could not be routed along that edge to a vertex, $v$, in $R_i$ which $e$ is on the unique path to in $T_i$ or $T^*_i$.

@@

Thus we observe that $f(T_i,x) = 2^i |T_i|$ and $f(T^*_i, x^*) = 2^i |T^*_i|$. Now in the first iteration (iteration $\delta$), $T_\delta$ is computed as an $O(1)$-approximation of $T^*_\delta$. Thus
$$2^\delta |T_\delta| = f(T_\delta, x) \leq O(1) f(T^*_\delta, x^*) = 2^\delta O(1) |T^*_\delta|.$$
Now consider the sub-tree that forms by taking the iterations of $$\bigcup_{j=\delta}^i T^*_j$$
and contracting $T_\delta \dots T_{i+1}$ into $r$. If any cycles are formed then do the following: while there still exists a cycle $C$ take the edge $e = \text{argmin}_{e\in C} x^*_e$ and remove it. Call the resulting tree $S^*$.
We claim that $(S^*,x^*)$ is feasible on $G_i$ for the nodes in $V(G_i) \cap R_i$, the graph which results from contracting $T_\delta, \dots, T_{i+1}$ into $r$. Since trimming cycles preserves connectivity, and we contracted edges into $r$, all $rv$-paths in $V(G_i) \cap R_i$ are preserved in $S^*$. Hence for all $v \in V(G_i) \cap R_i$ there exists an $rv$-path in $S^*$. Now we claim that the capacities on any $rv$-path still admit a feasible flow after converting to $S^*$. Observe that contraction will not decrease the minimum capacity. Further observe that deleted edges had minimum capacity on a cycle, so any flow that used a deleted edge $uv$ could send the same flow through the $uv$-path remaining in $S^*$ instead. Hence the flows on $S^*$ under capacities $x^*$ still meet the demands in $v \in V(G_i) \cap R_i$. Therefore $(S^*, x^*)$ feasible on $G_i$. Thus by our choice of $T_i$,
$$f(T_i, x) \leq O(1) f(S^*,x^*).$$
That is $$2^i |T_i| \leq 2^i O(1)|S^*| \leq 2^iO(1) (\sum_{j=\delta}^i |T^*_j|).$$
Summing these inequalities over $i$ we obtain:

@@

f(T,x) &= \sum_{i=1}^\delta 2^i|T_i| \\
&\leq \sum_{i=1}^\delta 2^i O(1)(\sum_{j=\delta}^i |T^*_j|)\\
&= O(1) \sum_{i=1}^\delta \sum_{j=i}^\delta 2^i|T^*_j| \\
&= O(1) \sum_{i=1}^\delta (\sum_{j=1}^{i}(2^j) |T^*_j| &\text{reindexing, and collecting coefficients on each $T^*_j$} \\
&= O(1) \sum_{i=1}^\delta 2^{i+1} |T^*_i| \\
&= 2O(1) \sum_{i=1}^\delta 2^i |T^*_i| \\
&= 2O(1) \sum_{i=1}^\delta f(T^*_i, x^*) \\
&= O(1) f(T^*, x^*).

@@

Thus the $O(1)$ approximation factor holds, and hence the given algorithm is an $O(1)$-approximation algorithm.$\blacksquare$
